---
title: "Untitled"
author: "Juliet"
date: "8/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(palmerpenguins)
library(tidyverse)
```

```{r}
# create a for loop that returns your total time for any two consecutive quarter miles

quarter_splits <- c(1.0, 1.1, 1.2, 1.1, 1.4, 1.5, 1.6, 1.4)

quarter_mi_consec <- vector(mode = "numeric", length = length(quarter_splits) - 1)

for (i in 1:(length(quarter_splits) - 1)) {
  quarter_mi_consec[i] <- quarter_splits[i] + quarter_splits[i + 1]
}

quarter_mi_consec

#use -1 in the seq_along(quarter_splits) - 1 because you cannot calc a value for the last i because there is no consec value, so there should only be 7 values returned

#can write with print function instead of naming it / storing it as quarter_mi_consec

# the following is the same thing with an intermediate object name:

quarter_mi_consec <- vector(mode = "numeric", length = length(quarter_splits) - 1)

for (i in 1:(length(quarter_splits) - 1)) {
  half_mile <- quarter_splits[i] + quarter_splits[i + 1]
  quarter_mi_consec[i] <- half_mile
}

quarter_mi_consec

#seq_along does not work well here because you want R to start at 1, not 0, and seq_along would output the first value as 0

quarter_mi_consec <- vector(mode = "numeric", length = length(quarter_splits) - 1)

for (i in (seq_along(quarter_splits) - 1)) {
  half_mile <- quarter_splits[i] + quarter_splits[i + 1]
  quarter_mi_consec[i] <- half_mile
}

quarter_mi_consec

```

# Basic Function Structure Practice

```{r}
bird_dog_sum <- function(bird, dog) {
  
  if (bird < 10) {
    message("cool little bird")
  }
  
  if (bird > 100) {
    error("birds cannot live that long")
  }
  
  if (dog < 0) {
    warning("can a dog be less than 1 lb?")
  }
  
  add_bird_dog <- bird + dog
  return(add_bird_dog)
}

#add conditions so your function will give warning or errors or messages for the coder. add this condition before the function process in the {} because you want this error to be spit out as soon as the coder initially enters the values


bird100_dog5 <- bird_dog_sum(bird = 8, dog = 5)

bird100_dog5

# needs numeric: bird_dog_sum(bird = "hawk", dog = 10) returns error


```

```{r}
say_hello <- function(friend, weekday) {
  paste("Hello", friend, "have a great", weekday)
}

say_hello(friend = "Teddy", weekday = "Thursday")

```

## Print() vs return() example

```{r}
find_mult <- function(a, b) {
  my_multiplication <- a * b
  print(my_multiplication)
}

find_mult(a = 10, b = 6)
find_mult(a = 10, b = 6) + 12
#python would not have allowed you to add 12 outside the function, python needs you the RETURN not just print if you want to do subsequent steps to the value
```

## Functions with conditions: have your function change based on the user's inputs, like depending on what species you pull data from

```{r}
animal_age <- function(animal, age) {
  
  if (!animal %in% c("dog", "goat")) {
    stop("unknown species entered")
  }
  
  if (animal == "dog") {
    creature_age <- age * 7
  }
  
  else if (animal == "goat") {
    creature_age <- age * 4.7
  }
  
  return(creature_age)
}

animal_age(animal = "goat", age = 13)

# now try to break this function, like another animal name or if the number is written as "ten"

#animal_age(animal = "zebra", age = 5)

#the ! before animal in the stop / error means NOT in that list of options


```



```{r}
# keep obs of adelie or gentoo

adelie_gentoo <- penguins %>% 
  filter(species %in% c("Adelie", "Gentoo"))

#R goes to the species col, then tests every row one by one, and the %in% asks if it matches adelie or gentoo and keeps it if so

# alternative way:

adelie_gentoo_wrong <- penguins %>% 
  filter(species == "Adelie" | species == "Gentoo")

#but this becomes tedious and long code, so it is more efficient to use the previous way

#common mistake:

adelie_gentoo_wrong <- penguins %>% 
  filter(species == c("Adelie", "Gentoo"))

# this is wrong because it does not detect every obs of adelie or gentoo, it is looking for an exact match, it asks if row 1 is adelie, then asks if row 2 is gentoo, then goes to row 3 and asks if it is adelie, it is alternating this vector until it reaches the end

# add a warning to this fucntion (if you wrote it) so a user can be warned that this function is probs not doing what you want it to do!


```


# combine functions and for loops!
```{r}
#take the mean of every column in a dataset just by entering the data frame

df_mean <- function(df) {
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]], na.rm = TRUE)
    print(col_mean)
  }
}

#test the function with mtcars

df_mean(df = mtcars)

# make the function return more specifically what column you are taking the mean of
# colnames(df[i]) will return that column name as a string

df_mean <- function(df) {
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]])
    print(paste("The mean of column", colnames(df[i]), "is", col_mean))
  }
}

# or:

df_mean <- function(df) {
  for (i in 1:ncol(df)) {
    col_mean <- mean(df[[i]])
    col_name <- colnames(df[i])
    print(paste("The mean of column", col_name, "is", col_mean))
  }
}

df_mean(df = mtcars)

```


```{r}
# make a subset of only structural size variables from the penguins dataset for chinstrap

chinstrap_sizes <- penguins %>% 
  filter(species == "Chinstrap") %>% 
  select(bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g)

# filter is for filtering rows, sleect is for subsetting columns

df_mean(df = chinstrap_sizes)
```


## Condition to check the class of an input

```{r}
lunch <- function(tacos, price) {
  
  if (is.numeric(tacos) == FALSE) {
    stop("you must input a number of tacos")
  }
  
  total_price <- tacos * price
  return(total_price)
}

lunch(tacos = 4, price = 2.50)
```

## loop example: use values from a data frame to calc something based on user input

```{r}
# working with different sites: mountain, prairie, desert, beach, each with depth and slope (cols) for each of these
# write a function that will automatically pull parameter functions from a data frame
# calc the groundwater transport rate

gw_rate <- function(site) {
  
  gw_depth_slope <- data.frame(sitename = c("mountain", "prairie", "beach", "desert"),
                               depth = c(32, 41, 63, 2),
                               slope = c(11.2, 0.4, 0.8, 2.6))
                               
  site_select <- dplyr::filter(gw_depth_slope, sitename == site)
  
  transport_rate <- 1.4 * site_select$slope + 3.6 * site_select$depth
  
  return(transport_rate)
}

# alternative for site_select line in the gw_rate function:   gw_depth_slope %>% filter(sitename == site)

gw_rate(site = "beach")

```

## switch function

```{r}
gw_rate_switch <- function(site) {
  
  gw_slope_depth <- switch(site,
                           "mountain" = c(32, 11.2),
                           "prairie" = c(41, 0.4),
                           "beach" = c(63, 0.8),
                           "desert" = c(2, 2.6))
  
  transport_rate <- 1.4 * gw_slope_depth[2] + 3.6 * gw_slope_depth[1]
  
  return(transport_rate)
}

gw_rate_switch(site = "beach")
```

# Logistic growth function

```{r}
  
logistic_growth <- function(K, N0, r, time) {
  Nt <- K / (1 +((K - N0) / N0) * exp(-r * time))
  print(Nt)
}

logistic_growth(K= 5000, N0 = 200, r = 0.43, time = 6)

# create a time sequence

time_vec <- seq(from = 0, to = 35, by = 0.1)
bacterial_growth <- logistic_growth(K = 5000, N0 = 200, r = 0.43, time = time_vec)

# in order to graph these against each other, you need to put them in the same dataframe

bacterial_df <- data.frame(time_vec, bacterial_growth)

ggplot(data = bacterial_df, aes(x = time_vec, y = bacterial_growth)) +
  geom_line(aes(color = time_vec))

```


```{r}
r_seq <- seq(from = 0.20, to = 0.40, by = 0.01)
time_vec <- seq(from = 0, to = 35, by = 0.1)

# create an output matrix
# (rows, col)

out_matrix <- matrix(ncol = length(r_seq), nrow = length(time_vec))

# N0 = 100, K = 6000

#write a nested for loop that uses my function and sends the output value for each combo into an output matrix in the order we discussed

for (i in seq_along(r_seq)) {
  for (j in seq_along(time_vec)) {
    pop <- logistic_growth(K = 6000, N0 = 100, r = r_seq[i], time = time_vec[j])
    out_matrix[j, i] <- pop
  }
}

#out_matrix

#the order of inner & outer loops doesn't matter, but J, i must match out_matrix order (ncol =, nrow =) which must correspond and therefore populate the matrix by row then col or col then row

# wrangling

out_df <- data.frame(time = time_vec, out_matrix)

colnames(out_df) <- c("time", paste0("gr_", r_seq))

# make long form (tidy version) of this data
out_df_long <- out_df %>% 
  pivot_longer(-time, names_to = "growth_rate", values_to = "population")


# make a line graph with pop on Y over time on x with color of line dependent on the growth rate

ggplot(data = out_df_long, aes(x = time, y = population)) +
  geom_line(aes(color = growth_rate))

```







